---
title: "hw4.0_tidy_data_and_joins"
author: "Xuan Lin"
date: '2019-10-07'
output: 
  html_document:
    keep_md: true
    toc: true
---

```{r warning=FALSE}
library(gapminder)
library(tidyverse)
```

## Exercise 1: Univariate Data Reshaping (Option 1)

### _1.1_

__Description__

We are interested in the life exectancies in Asian countries from year 1952 to 2007. The original _gapminder_ dataset presents each year and contry in seperate rows which makes comparison of life expectancy of different countries in the same year somehow difficult. The following table filtered out Asian countries and re-organized the columns so that direct comparison of life expectancy in the same year can be easily achieved.

```{r warning=FALSE}
gapminder %>% 
  filter(continent == "Asia") %>% 
  pivot_wider(id_cols = year, names_from = country, values_from = lifeExp) %>% 
  DT::datatable()
```

__Table 1 Life expectancy in Asia (wide format)__

### _1.2_

__Description__

We then compare life expectancy in China and Japan by plotting them together. The red and blue lines show life expectancy in China and Japan, respectively.

```{r warning=FALSE}
gapminder %>% 
  filter(continent == "Asia") %>% 
  pivot_wider(id_cols = year, names_from = country, values_from = lifeExp) %>% 
  ggplot(aes(x=year)) +
  geom_line(aes(y= China, color = "red")) +
  geom_line(aes(y= Japan, color= "blue")) +
  scale_y_continuous("Life Expectancy") +
  scale_color_discrete("Country", labels = c("Japan", "China")) +
  scale_x_continuous("Year") +
  theme_classic()
```

__Fig.1 Life expectancy in China and Japan (1952-2007)__

### _1.3_

__Description__

We finally re-organize the wide table made in the first step back to the original format.

```{r warning=FALSE}
gapminder %>% 
  filter(continent == "Asia") %>% 
  pivot_wider(id_cols = year, names_from = country, values_from = lifeExp) %>% 
  pivot_longer(cols = -year, names_to = "country", values_to = "lifeExp") %>% 
  DT::datatable()
```

__Table 2 Life expectancy in Asia (original)__

## Exercise 2: Multivariate Data Reshaping (Option 1)

### _2.1_

__Description__

Except life expectancy, we are also interested in GDP per capita in Asian countries. Therefore, we make the following wide table of life expectancy and GDP per capita in Asian counties from 1952 to 2007.

```{r warning=FALSE}
gapminder %>% 
  filter(continent == "Asia") %>%
  pivot_wider(id_cols = year, 
              names_from = country,
              names_sep = "_",
              values_from = c(lifeExp, gdpPercap)) %>% 
  DT::datatable()
```

__Table 3 Life expectancy and GDP per capita in Asian countries (wide format)__

### 2.2

__Description__

The wide-formatted table is then reshaped to its original format. To do this, we first convert the wide formatted table to a longer table having the combination of each year and life expectancy / GDP per capita for all countries as a single row, then reshape it to the longest table having the combination of each year, life expectancy/ GDP per capita and country as a sigle row, finally widen the longest table to the original format.

```{r warning=FALSE}
gapminder %>% 
  filter(continent == "Asia") %>%
  pivot_wider(id_cols = year, 
              names_from = country,
              names_sep = "_",
              values_from = c(lifeExp, gdpPercap)) %>% 
  pivot_longer(cols = -year,
               names_to = c("index",".value"),
               names_sep = "_") %>% 
  pivot_longer(cols = -c(year,index), names_to = "country", values_to = "measurements") %>% 
  pivot_wider(id_cols = c(year, country), names_from = index, values_from = "measurements") %>% 
  DT::datatable()
```

__Table 4 Life expectancy and GDP per capita in Asian countries (original format)__

## Exercise 3: Table Joins
```{r warning=FALSE}
guest <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/attend.csv")
email <- read_csv("https://raw.githubusercontent.com/STAT545-UBC/Classroom/master/data/wedding/emails.csv")
```

### _3.1_

__Descrpition__

For each guest in the _guestlist_ (guest tibble), add a column for email address, which can be found in the _email_ tibble. 

We first separate the names in the original _email_ tibble and save it as a new tibble **email_address**, then using *left_join* to add the email address to the guest list.

```{r warning=FALSE}
email_address <- email %>% 
  separate_rows(guest, sep = ", ") %>% 
  rename(name = guest)

guest %>% 
  left_join(email_address, by = "name") %>% 
  knitr::kable()

```

### _3.2_

__Descrpition__

Who do we have emails for, yet are not on the guestlist?

```{r warning=FALSE}
anti_join(email_address, guest) %>% 
  knitr::kable()
```


### _3.3_

__Descrpition__

Make a guestlist that includes everyone we have emails for (in addition to those on the original guestlist).

```{r warning=FALSE}
full_join(guest, email_address) %>% 
  knitr::kable()
```

