---
title: "hw2.0-Gapminder and dplyr"
author: "Xuan Lin"
date: '2019-09-21'
output: 
  html_document:
    keep_md: true
    toc: true
---

## Setup
```{r echo=FALSE, warning=FALSE}
library(gapminder)
library(tidyverse)
```

## _Exercise 1_ Basic dplyr
### 1.1


__Table 1.1 Life expectancy, population and GDP in Canada, China and Japan (1970's)__

```{r echo=FALSE}
filter(gapminder, year== 1972 | year== 1977, country== "China" | country=="Canada" | country=="Japan") %>% 
knitr::kable()
```



### 1.2

__Table 1.2 GDP in Canada, China and Japan (1970's)__


```{r echo=FALSE}
filter(gapminder, year== 1972 | year== 1977, country== "China" | country=="Canada" | country=="Japan") %>% 
  select(country, gdpPercap) %>% 
  knitr::kable()
```


### 1.3

__Table 1.3 Countries experienced a drop in life expectancy (1952-2007)__

```{r echo=FALSE}
gapminder %>% 
  group_by(country) %>% 
  mutate(increaseInlifeExp = c(0, round(diff(lifeExp, 1), digits = 1))) %>% 
  filter(increaseInlifeExp < 0) %>% 
  DT::datatable()
```


### 1.4

__Table 1.4 Maximum GDP experienced by each country (1952-2007)__

```{r echo=FALSE}
gapminder %>% 
  group_by(country) %>% 
  filter(gdpPercap == max(gdpPercap)) %>% 
  DT::datatable()
```


### 1.5


```{r echo=FALSE}
gapminder %>%
  filter(country == "Canada") %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point() +
  scale_y_log10("GDP per capita", labels = scales::dollar_format()) +
  xlab("Life Expectancy")
```

__Fig 1.1 Relationship between life expectancy and GDP in Canada (1952-2007)__

## _Exercise 2_ Explore individual variables with dplyr

__Table 2.1 Minimum and maximum life expetancy each year__

```{r echo=FALSE}
gapminder %>%
  group_by(year) %>% 
  filter(lifeExp == max(lifeExp) | lifeExp == min(lifeExp)) %>% 
  select(year, lifeExp) %>% 
  arrange(year, lifeExp) %>% 
  knitr::kable()
```




```{r echo=FALSE}
gapminder %>%
  ggplot(aes(year, lifeExp)) +
  geom_boxplot(aes(group = year)) +
  xlab("Year") +
  ylab("Life Expectancy")
```

__Fig 2.1 Boxplot of life expectancy each year (1952-2007)__

## _Exercise 3_ Explore various plot types
### Scatterplot


```{r echo=FALSE}
gapminder %>%
  filter(country == "China") %>% 
  ggplot(aes(pop, gdpPercap)) +
  geom_point(color = "black", fill = "Navy", shape = 22, alpha = 0.5, size = 6) +
  scale_x_continuous ("Population", labels = scales::scientific_format(digits = 2, scale = 1)) +
  scale_y_continuous ("GDP per capita", labels = scales::dollar_format())
```


__Fig 3.1 Relationship between Chinese population and GDP per capita (1952 - 2007)__


### Boxplot

```{r echo=FALSE}
gapminder %>%
  filter(year == 2007) %>% 
  mutate(popB = round(log10(pop), digit=0)) %>% 
  group_by(popB) %>% 
  ggplot(aes(popB, gdpPercap)) +
  geom_boxplot(aes(group=popB)) +
  scale_x_continuous("Population", labels = scales::math_format(expr = 10^.x)) +
  scale_y_log10 ("GDP per capita", labels = scales::dollar_format())
```


__Fig 3.2 Difference in GDP per capita with different range of population (2007)__